---
title: "cate-causal-analysis"
author: "Matt Rosinski"
date: "2023-04-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
set.seed(42)

# Number of samples
n <- 5000

# Covariate: age
age <- rnorm(n, mean = 30, sd = 10)

# Treatment variable: job training (random assignment)
treatment <- rbinom(n, size = 1, prob = 0.5)

# True treatment effect (varying with age)
# Here's how the treatment effect varies with age:
# 
# We use the cumulative distribution function of a normal distribution (pnorm) 
# with a mean of 40 and a standard deviation of 10 to create a weight that 
# depends on an individual's age.
# We subtract this weight from 1, so that the treatment effect is larger for 
# younger individuals (with ages less than 40) and smaller for older individuals 
# (with ages greater than 40).
# We multiply this adjusted weight by 500 to scale the treatment effect to a 
# more realistic range.
true_cate <- 500 * (1 - pnorm(age, mean = 40, sd = 10))

# Outcome variable: income
noise <- rnorm(n, mean = 0, sd = 500)
income <- 2000 + 50 * age + treatment * true_cate + noise

# Create the data frame
data <- tibble(age = age, treatment = treatment, income = income)
```


```{r}
model <- lm(income ~ treatment * age, data = data)
summary(model)

# The estimated coefficients for treatment, age, and 
# treatment:age represent the treatment effect, the effect of age on income, 
# and the heterogeneous treatment effect (CATE) across different ages, respectively.

```

```{r}
# In this case, the model assumes that the effect of the treatment is the 
# same for all individuals, regardless of their age. This model only 
# considers the main effects of treatment and age on income.
model_lm <- lm(income ~ treatment + age, data = data)
summary(model_lm)
```

```{r}
# when you fit the model lm(income ~ treatment * age, data = data), you're 
# including the interaction term between treatment and age. This model allows 
# for the treatment effect to vary across different ages, capturing the 
# heterogeneous treatment effect.
```

# Compare goodness of fit
```{r}
# Model without interaction
model_no_interaction <- lm(income ~ treatment + age, data = data)

# Model with interaction
model_interaction <- lm(income ~ treatment * age, data = data)

```

# Calculate AIC and BIC
Akaike Information Criterion (AIC) and Bayesian Information Criterion (BIC)
```{r}
aic_no_interaction <- AIC(model_no_interaction)
bic_no_interaction <- BIC(model_no_interaction)

aic_interaction <- AIC(model_interaction)
bic_interaction <- BIC(model_interaction)

```

# Compare AIC and BIC values
If the AIC and/or BIC values for the model with interaction are lower 
than the ones for the model without interaction, it suggests that the 
model with interaction is superior, taking both goodness of fit and model 
complexity into account.
```{r}
# Compare AIC values
cat("AIC without interaction:", aic_no_interaction, "\n")
cat("AIC with interaction:", aic_interaction, "\n")

# Compare BIC values
cat("BIC without interaction:", bic_no_interaction, "\n")
cat("BIC with interaction:", bic_interaction, "\n")

```
# Exploratory data analysis
```{r}
library(ggplot2)

data %>%
  ggplot(aes(x = age, y = income, color = as.factor(treatment))) +
  geom_point(alpha = 0.5) +
  theme_minimal() +
  labs(x = "Age", y = "Income", color = "Treatment") +
  ggtitle("Scatterplot of Income by Age and Treatment")


```
Plot the model predictions against the true income values.
```{r}
data$predicted_no_interaction <- predict(model_no_interaction, data)
data$predicted_interaction <- predict(model_interaction, data)

# Plot model predictions against true income values
data %>%
  ggplot() +
  geom_point(aes(x = income, y = predicted_no_interaction, color = "No Interaction"), alpha = 0.5) +
  geom_point(aes(x = income, y = predicted_interaction, color = "Interaction"), alpha = 0.5) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black") +
  theme_minimal() +
  labs(x = "True Income", y = "Predicted Income", color = "Model") +
  ggtitle("Model Predictions vs. True Income")

```

# Compare residuals
```{r}
data$residuals_no_interaction <- residuals(model_no_interaction)
data$residuals_interaction <- residuals(model_interaction)

# Residuals vs. Age
data %>%
  ggplot() +
  geom_point(aes(x = age, y = residuals_no_interaction, color = "No Interaction"), alpha = 0.5) +
  geom_point(aes(x = age, y = residuals_interaction, color = "Interaction"), alpha = 0.5) +
  theme_minimal() +
  labs(x = "Age", y = "Residuals", color = "Model") +
  ggtitle("Model Residuals vs. Age")

```

```{r}
# Add a new column to the data for model type
data_long <- data %>%
    select(age, residuals_no_interaction, residuals_interaction) %>%
    pivot_longer(cols = 2:3, names_to = "Model", values_to = "Residuals")

# Residuals vs. Age
ggplot(data_long, aes(x = age, y = Residuals, color = Model)) +
    geom_point(alpha = 0.5) +
    facet_wrap(~ Model, scales = "free_y") +
    theme_minimal() +
    theme(legend.position = "bottom") +
    labs(title = "Model Residuals vs. Age", x = "Age", y = "Residuals", color = "Model") 

```
# Perform k-folds cross validation
```{r}
library(tidymodels)
library(doParallel)
library(foreach)
library(tictoc)

set.seed(42)
data_split <- initial_split(data, prop = 0.8)
data_train <- training(data_split)
data_test <- testing(data_split)

```

```{r}
lm_spec_no_interaction <- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

lm_spec_interaction <- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

```

```{r}
recipe_no_interaction <- recipe(income ~ treatment + age, data = data)

recipe_interaction <- recipe(income ~ treatment + age, data = data) %>%
  step_interact(~ treatment * age)

```

```{r}
set.seed(42)
registerDoParallel(cores = parallel::detectCores())

# Need to check for data leakage
# test and train split


# Define k-fold cross-validation settings
cv_folds <- vfold_cv(data_train, v = 10)

# Fit the models using cross-validation in parallel
fit_resamples_control <- control_resamples(save_pred = TRUE, verbose = FALSE)
```


```{r}
tic()
# Fit the model without interaction using cross-validation
cv_no_interaction <- workflow() %>%
  add_model(lm_spec_no_interaction) %>%
  add_recipe(recipe_no_interaction) %>%
  fit_resamples(cv_folds, control = fit_resamples_control)

# Fit the model with interaction using cross-validation
cv_interaction <- workflow() %>%
  add_model(lm_spec_interaction) %>%
  add_recipe(recipe_interaction) %>%
  fit_resamples(cv_folds, control = fit_resamples_control)
toc()

```


```{r}
tic()
# Fit the model without interaction using cross-validation in parallel
cv_no_interaction <- foreach(i = seq_along(cv_folds$splits), 
                             .combine = c, 
                             .packages = "tidymodels") %dopar% {
  split <- cv_folds$splits[[i]]
  data_train <- analysis(split)
  data_test <- assessment(split)
  
  fit_no_interaction <- lm_spec_no_interaction %>%
    fit(income ~ treatment + age, data = data_train)
  
  predict(fit_no_interaction, data_test) %>%
    bind_cols(data_test, .)
}

# Fit the model with interaction using cross-validation in parallel
cv_interaction <- foreach(i = seq_along(cv_folds$splits), 
                          .combine = c, 
                          .packages = "tidymodels") %dopar% {
  split <- cv_folds$splits[[i]]
  data_train <- analysis(split)
  data_test <- assessment(split)
  
  fit_interaction <- lm_spec_interaction %>%
    fit(income ~ treatment * age, data = data_train)
  
  predict(fit_interaction, data_test) %>%
    bind_cols(data_test, .)
}
toc()

```
# The model with interaction has a lower RMSE
```{r}
# Collect performance metrics
performance_no_interaction <- cv_no_interaction %>% collect_predictions()
performance_interaction <- cv_interaction %>% collect_predictions()

# Calculate mean squared errors
mse_no_interaction <- mean((performance_no_interaction$income - performance_no_interaction$.pred)^2)
mse_interaction <- mean((performance_interaction$income - performance_interaction$.pred)^2)

cat("Mean RMSE without interaction (training data):", sqrt(mse_no_interaction), "\n")
cat("Mean RMSE with interaction (training data):", sqrt(mse_interaction), "\n")

```
# Final retrain and test
```{r}
# Train final models on the entire training dataset
final_model_no_interaction <- lm_spec_no_interaction %>% fit(income ~ treatment + age, data = data_train)
final_model_interaction <- lm_spec_interaction %>% fit(income ~ treatment * age, data = data_train)

# Make predictions on the test dataset
predictions_no_interaction <- predict(final_model_no_interaction, data_test)
predictions_interaction <- predict(final_model_interaction, data_test)

# Calculate mean squared errors for the test dataset
mse_no_interaction_test <- mean((data_test$income - predictions_no_interaction$.pred)^2)
mse_interaction_test <- mean((data_test$income - predictions_interaction$.pred)^2)

cat("Mean RMSE without interaction (test data):", sqrt(mse_no_interaction_test), "\n")
cat("Mean RMSE with interaction (test data):", sqrt(mse_interaction_test), "\n")

```

