---
title: "object oriented programming"
author: "Matt Rosinski"
date: '2022-04-30'
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{python}
class Employee:
  
  def set_name(self, new_name):
    self.name = new_name
  
  # Add set_salary() method
  def set_salary(self, new_salary):
    self.salary = new_salary

```

```{python}
# Create an object emp of class Employee  
emp = Employee()

# Use set_name to set the name of emp to 'Korel Rossi'
emp.set_name('Korel Rossi')

# Set the salary of emp to 50000
emp.set_salary(50000)

# We can access the attributes like this
print(emp.name)
print(emp.salary)
```

```{r}
library(R6)

?R6::R6Class()

Employee = 
    R6Class("Employee",
            # Store data you don't want the user to access in the private element
               private = list(
                   name = 'Blank',
                   salary = 0
               ),
            # Store functions you want the user to access in the active or public element
               public = list(
                   set_name =  function(new_name) {
                       private$name <-  new_name},
                   set_salary = function(new_salary) {
                       private$salary <- new_salary},
                   get_name = function() private$name,
                   get_salary = function() private$salary,
                   initialize = function(name, salary) {
                       if(!missing(name)) {
                           private$name <- name}
                       if(!missing(salary)) {
                           private$salary <- salary}}
               ))
```

# First we forget to specify the name and salary
```{r}
emp <- Employee$new()
```

# So the name and salary are blank
```{r}
emp$get_name()
emp$get_salary()

```

# We can update the object with the name and salary after creation
```{r}
emp$set_name('Korel Rossi')
emp$set_salary('50000')
emp$get_name()
emp$get_salary()
```

```{r}
library(assertive)

Employee = 
    R6Class("Employee",
            # Store data you don't want the user to access in the private element
               private = list(
                   ..name = 'Blank',
                   ..salary = 0
               ),
            # Store functions you want the user to access in the active or public element
               # public = list(
               #     initialize = function(name, salary) {
               #         if(!missing(name)) {
               #             private$name <- name}
               #         if(!missing(salary)) {
               #             private$salary <- salary}}
               # ),
            # Active binding allow for restricted access to the private data
               active = list(
                   name =  function(value) {
                       if(missing(value)) {
                           private$..name} else {
                               assert_is_a_string(value)
                               private$..name <- value
                               }
                       },
                   salary = function(value) {
                       if(missing(value)) {
                           private$..salary} else {
                               assert_is_a_number(value)
                               private$..salary <- value
                           }
                   }
               ))
```

```{r}
emp = Employee$new()
```

```{r}
emp$name
emp$salary
```

```{r}
emp$name <- 'Korel Rossi'
emp$salary <- 50000
```

```{r}
emp$name
emp$salary
```

```{r}
Person <- R6Class("Person",
  public = list(
    name = NULL,
    hair = NULL,
    initialize = function(name = NA, hair = NA) {
      self$name <- name
      self$hair <- hair
      # self$greet()
    }
    # set_hair = function(val) {
    #   self$hair <- val
    # },
    # greet = function() {
    #   cat(paste0("Hello, my name is ", self$name, ".\n"))
    # }
  )
)
```

```{r}
pp <- Person$new()
```

```{r}
pp$name <- 'Bob'
pp$hair <- 'Blonde'
```

```{r}
pp$name
pp$hair
```

# Define a class in python and return monthly salary attribute.
```{python}
class Employee:
    def set_name(self, new_name):
        self.name = new_name
    
    def set_salary(self, new_salary):
        self.salary = new_salary
        
    def give_raise(self, amount):
        self.salary = self.salary + amount
    
    def monthly_salary(self):
        return self.salary/12

    
emp = Employee()
emp.set_name('Korel Rossi')
emp.set_salary(50000)

# Get monthly salary of emp and assign to mon_sal
mon_sal = emp.monthly_salary()

# Print mon_sal
print(mon_sal)
```

# Using conditions to check entries
```{python}
class Employee:
  
    def __init__(self, name, salary=0):
        self.name = name
        # Modify code below to check if salary is positive
        if salary > 0:
            self.salary = salary
        else:
            self.salary = 0
            print('Invalid salary!')
   
   # ...Other methods omitted for brevity ...
      
emp = Employee("Korel Rossi", -1000)
print(emp.name)
print(emp.salary)
```

# Adding more attributes to class initialisation
```{python}
# Import datetime from datetime
from datetime import datetime

class Employee:
    
    def __init__(self, name, hire_date, salary=0):
        self.name = name
        if salary > 0:
          self.salary = salary
        else:
          self.salary = 0
          print("Invalid salary!")
          
        # Add the hire_date attribute and set it to today's date
        self.hire_date = datetime.today()
        
   # ...Other methods omitted for brevity ...
      
emp = Employee("Korel Rossi", -1000)
print(emp.name)
print(emp.salary)
```

# Define a pretend visualisation class
```{python}
from numpy import sqrt
class Point:

    def __init__(self, x=0.0, y=0.0):
        self.x = x 
        self.y = y 
    
    def distance_to_origin(self):
        return sqrt(self.x ** 2 + self.y **2)
    
    def reflect(self, axis):
        if axis == 'x':
            self.y = -self.y
        elif axis == 'y':
            self.x = -self.x
        else:
            print("Enter either 'x' or 'y' for axis to reflect")

pt = Point(x=3.0)
pt.reflect("y")
print((pt.x, pt.y))
pt.y = 4.0
print(pt.distance_to_origin())

```
# Class level attributes
```{python}
# Create a Player class
class Player:
    MAX_POSITION = 10
    def __init__(self, position=0):
        self.position = position


# Print Player.MAX_POSITION       
print(Player.MAX_POSITION)

# Create a player p and print its MAX_POSITITON
p = Player()
print(p.MAX_POSITION)
```

# Very sensitive to spacing
```{python}
class Player:
    MAX_POSITION = 10
    
    def __init__(self):
        self.position = 0
        
    def move(self, steps):
        if self.position + steps < Player.MAX_POSITION:
            self.position = self.position + steps
        else:
            self.position = Player.MAX_POSITION
            
    def draw(self):
        drawing = "-" * self.position + "|" +"-"*(Player.MAX_POSITION - self.position)
        print(drawing)

p = Player()

p.draw()
p.move(4); p.draw()
p.move(5); p.draw()
p.move(3); p.draw()
```

```{python}
class BetterDate:    
    # Constructor
    def __init__(self, year, month, day):
      # Recall that Python allows multiple variable assignments in one line
      self.year, self.month, self.day = year, month, day
    
    # Define a class method from_str
    @classmethod
    def from_str(cls, datestr):
        # Split the string at "-" and convert each part to integer
        parts = datestr.split("-")
        year, month, day = int(parts[0]), int(parts[1]), int(parts[2])
        # Return the class instance
        return cls(year, month, day)
        
bd = BetterDate.from_str('2020-04-30')   
print(bd.year)
print(bd.month)
print(bd.day)
```
# Class method definition
```{python}
# import datetime from datetime
from datetime import datetime

class BetterDate:
    def __init__(self, year, month, day):
      self.year, self.month, self.day = year, month, day
      
    @classmethod
    def from_str(cls, datestr):
      year, month, day = map(int, datestr.split("-"))
      return cls(year, month, day)
      
    # Define a class method from_datetime accepting a datetime object
    @classmethod
    def from_datetime(cls,datetime):
      year = datetime.year
      month = datetime.month
      day = datetime.day
      return cls(year, month, day)

# You should be able to run the code below with no errors: 
today = datetime.today()     
bd = BetterDate.from_datetime(today)   
print(bd.year)
print(bd.month)
print(bd.day)
```

# Create a subclass
```{python}
class Employee:
    MIN_SALARY = 30000
    
    def __init__(self, name, salary=MIN_SALARY):
        self.name = name
        if salary >= Employee.MIN_SALARY:
            self.salary = salary
        else:
            self.salary = Employee.MIN_SALARY
    def give_raise(self, amount):
        self.salary += amount      
        
# MODIFY Manager class and add a display method
class Manager(Employee):
    def __init__(self, name, salary):
        super().__init__(name, salary)
    
    def display(self):
        print(f'Manager {self.name}')
        return

mng = Manager("Debbie Lashko", 86500)
print(mng.name)

# Call mng.display()
mng.display()
```

```{python}
mng.MIN_SALARY
```

# If you can't explain it, you don't know it.  Master the fundamental principles.
# See what's there, what's simple and familiar