---
title: "exploring spotify api"
author: "Matt Rosinski"
date: "2023-03-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Future exploration - Youtube API
```{r}
# https://github.com/soodoku/tuber

```


# Spotify API
```{r}
# devtools::install_github('charlie86/spotifyr')

# Documentation
# https://www.rcharlie.com/spotifyr/
Sys.setenv(SPOTIFY_CLIENT_ID = 'getyourown')
Sys.setenv(SPOTIFY_CLIENT_SECRET = 'gotospotifywebapi')
Sys.setenv(SPOTIFY_REDIRECT_URI = "http://localhost:1410/")


```

# Spotifyr documentation
https://www.rdocumentation.org/packages/spotifyr/versions/2.2.4

```{r}
library(spotifyr)
library(tidyverse)
library(lubridate)
library(gt)

access_token <- get_spotify_access_token()
```

```{r}
beatles <- get_artist_audio_features('the beatles')

```

# What was The Beatles favourite key?
```{r}

beatles %>% 
    count(key_mode, sort = TRUE) %>% 
    head(5) %>% 
    gt()

beatles %>% glimpse()

```
```{r}
my_id <- 'loudfruit'
my_plists <- get_user_playlists(user_id = my_id)

my_plists %>% str()

my_plists %>% names()

my_plists %>%  View()

my_plists2 <- my_plists %>%
    filter(name %in% c('2023 Most Streamed', 'Most Streamed')) %>% 
    select(href) %>% pull()
    

```

# Get my recently played songs
```{r}

get_my_recently_played(limit = 10) %>%
    mutate(artist.name = map_chr(track.artists, function(x) x$name[1]),
           played_at = as_datetime(played_at)) %>%
    select(track.name, artist.name, track.album.name, played_at) %>%
    gt()


```


```{r}
get_my_playlists()

get_user_playlists("loudfruit")

res <- get_artist_top_tracks(id = "0TnOYISbd1XYRBk9myaseg")
res$artists[1]

res$name
res$album.name
res$album.artists
```

# Find the Spotify ID for an artist
```{r}
# Replace this with the name of the artist you are looking for
artist_name <- "Radiohead"

# Search for the artist
search_results <- search_artists(artist_name)

# Display the search results
search_results

spotify_id <- 
    spotifyr::search_spotify(q = artist_name, type = "artist") %>% 
    arrange(desc(popularity)) %>% 
    select(-c("genres", "href", "images")) %>% 
    select(id) %>% 
    slice_head(n = 1) %>% 
    pull()

spotify_id


```


# Get my followed artists
```{r}
get_my_followed_artists(
  limit = 20,
  after = NULL,
  authorization = get_spotify_authorization_code(),
  include_meta_info = FALSE
) %>% 
    select(name)



# Check authorisation code
# get_spotify_authorization_code(
#   client_id = Sys.getenv("SPOTIFY_CLIENT_ID"),
#   client_secret = Sys.getenv("SPOTIFY_CLIENT_SECRET"),
#   scope = scopes()
# )

```

# Get Spotify ID for playlists
```{r}

search_string <- "Australia"

spotify_playlists <- 
    spotifyr::search_spotify(q = search_string, type = "playlist") %>% 
    select(id, name, tracks.total, uri, external_urls.spotify, owner.id, owner.display_name) 

spotify_playlist_id <- spotify_playlists %>% 
    select(id) %>% 
    slice_head(n = 1) %>% 
    pull()

spotify_playlist_id


```

# Get User Playlists
```{r}

user_id <- "spotify"

user_playlists = get_user_playlists(
  user_id = user_id,
  limit = 20,
  offset = 0,
  authorization = get_spotify_authorization_code(),
  include_meta_info = FALSE
)

user_playlists %>%  glimpse()

user_playlists %>% 
    select(name, id, uri)

playlist_uris <- 
    user_playlists %>% 
    select(id) %>% 
    pull()

playlist_uris

```
# Get single playlist
```{r}

playlist_id <- sample(playlist_uris, size = 1)

playlist_df <- get_playlist(
  playlist_id = playlist_id,
  fields = NULL,
  market = NULL,
  authorization = get_spotify_access_token()
)

names(playlist_df)

tidy(playlist_df)

```
# Get my playlist
```{r}
myplaylists <- get_my_playlists()

myplaylists

```


# Get popularity and audio features of every song in a playlist
```{r}

playlist_song_features <- 
    get_playlist_audio_features(
        username = "spotify",
        playlist_uris = playlist_uris,
        authorization = get_spotify_access_token()
)

playlist_song_features %>% glimpse()

playlist_song_features %>% 
    select(track.popularity, 
           track.name, 
           track.artists, 
           track.album.name, 
           track.album.release_date, 
           track.id,
           danceability:tempo
           ) %>% 
    glimpse()

write_csv(playlist_song_features, file = "data/playlist_song_features.csv")

# playlist_song_features$track.artists

```
# Get album information
```{r}

albumn_ids <- sample(playlist_song_features$track.album.id, size = 20)

album_info_df <- get_albums(
  ids = albumn_ids,
  market = NULL,
  authorization = get_spotify_access_token(),
  include_meta_info = FALSE
)

names(album_info_df)

album_info_df$artists

album_info_df$name


```
# Search Spotify
```{r}
# search_spotify(
#   q,
#   type = c("album", "artist", "playlist", "track", "show", "episode"),
#   market = NULL,
#   limit = 20,
#   offset = 0,
#   include_external = NULL,
#   authorization = get_spotify_access_token(),
#   include_meta_info = FALSE
# )
search_string <- "Radiohead"

search_results <- search_spotify(
  q = search_string,
  type = c("artist"),
  authorization = get_spotify_access_token()
)

artist_id <- search_results %>% 
    arrange(desc(popularity)) %>% 
    select(id, name, uri, popularity, followers.total) %>% 
    slice_head(n = 1) %>% 
    select(id) %>% 
    pull()
```


# Get artist audio features
```{r}
search_string <- "Radiohead"

# Uses search_spotify() for precise selection of artist id
radiohead_audio_dt <- 
    get_artist_audio_features(
    artist = artist_id,
    include_groups = "album",
    return_closest_artist = TRUE,
    dedupe_albums = TRUE,
    market = NULL,
    authorization = get_spotify_access_token()
    )

radiohead_audio_dt %>% glimpse()

# Combine with track popularity using track_id

```
# Get artist top tracks by country

```{r}
radiohead_top_us_tracks <- get_artist_top_tracks(
  id = artist_id,
  market = "US",
  authorization = get_spotify_access_token(),
  include_meta_info = FALSE
)

radiohead_popular_us_df <- radiohead_top_us_tracks %>%  
    arrange(desc(popularity)) %>% 
    mutate(country = "US") %>% 
    select(name, id, popularity, country, uri, album.name)

radiohead_popular_au_df <- radiohead_top_us_tracks %>%  
    arrange(desc(popularity)) %>% 
    mutate(country = "AU") %>% 
    select(name, id, popularity, country, uri, album.name)

radiohead_tracks_us <- radiohead_popular_us_df %>% 
    left_join(radiohead_audio_dt, by = c("name" = "track_name")) %>% 
    select(-c("album_images", "track_preview_url", "type", "external_urls.spotify"))

nrow(radiohead_tracks_us)

radiohead_tracks_au <- radiohead_popular_au_df %>% 
    left_join(radiohead_audio_dt, by = c("name" = "track_name")) %>% 
    select(-c("album_images", "track_preview_url", "type", "external_urls.spotify"))

nrow(radiohead_tracks_au)

radiohead_tracks_us %>% 
    anti_join(radiohead_tracks_au, by = "name")

radiohead_tracks_au %>% 
    anti_join(radiohead_tracks_us, by = "name")

radiohead_tracks_us %>% 
    select(name, popularity, country) %>% 
    left_join(radiohead_tracks_au %>% 
    select(name, popularity, country), by = "name")
```

# Get track information including popularity

```{r}
track_id <- radiohead_top_us_tracks %>% select(id) %>% slice_head(n = 1) %>%  pull()

track_info <- get_track(id = track_id, market = NULL, authorization = get_spotify_access_token())

names(track_info)

track_info$popularity


```
# Get track popularity scores
```{r}
track_ids <- radiohead_top_us_tracks %>% select(id) %>% slice_head(n = 10) %>%  pull()

track_list_info <- 
    get_tracks(
    ids = track_ids,
    market = NULL,
    authorization = get_spotify_access_token(),
    include_meta_info = FALSE
)

popularity_score_df <- 
    track_list_info %>% 
    select(id, name, popularity, duration_ms, uri, album.name, album.release_date)

```



# Get my top artists - returning empty list()
```{r}
get_my_top_artists_or_tracks(
  type = "artists",
  limit = 20,
  offset = 0,
  time_range = "short_term",
  authorization = get_spotify_authorization_code(),
  include_meta_info = FALSE
)
```


```{r}
get_my_top_artists_or_tracks(
    type = 'artists', 
    time_range = 'short_term', 
    limit = 5) %>% 
    select(name, genres) %>% 
    rowwise %>% 
    mutate(genres = paste(genres, collapse = ', ')) %>% 
    ungroup %>% 
    gt()
```
```{r}


get_my_top_artists_or_tracks(type = 'artists', 
                             time_range = 'long_term', 
                             limit = 5) %>% 
    select(.data$name, .data$genres) %>% 
    rowwise %>% 
    mutate(genres = paste(.data$genres, collapse = ', ')) %>% 
    ungroup %>% 
    gt()

get_my_top_artists_or_tracks(type = 'tracks', 
                             time_range = 'short_term', 
                             limit = 5) %>% 
    mutate(artist.name = map_chr(artists, function(x) x$name[1])) %>% 
    select(name, artist.name, album.name) %>% 
    gt()
```

```{r}
library(dplyr)
library(gt)
library(purrr)

access_token <- get_spotify_access_token()

spotifyr::get_my_top_artists_or_tracks()

get_my_top(type = 'tracks', time_range = 'short_term', limit = 5) %>% 
    mutate(artist.name = map_chr(artists, function(x) x$name[1])) %>% 
    select(name, artist.name, album.name) %>% 
    gt()

```


```{r}
radiohead <- spotifyr::get_artist_audio_features("radiohead")

```

```{r}
radiohead %>% glimpse()

radiohead_id <- unique(radiohead$artist_id)

radiohead_top_tracks <- spotifyr::get_artist_top_tracks(id = radiohead_id)

radiohead_top_tracks %>% glimpse()

```

```{r}
joy <- get_artist_audio_features('joy division')
```


```{r}

joy %>% 
    arrange(-valence) %>% 
    select(track_name, valence) %>% 
    head(5) %>% 
    gt()

joy %>% 
    arrange(-valence) %>% 
    select(.data$track_name, .data$valence) %>% 
    head(5) %>% 
    gt()
```
```{r}
library(ggjoy)

ggplot(joy, aes(x = valence, y = album_name)) + 
    geom_joy() + 
    theme_joy() +
    ggtitle("Joyplot of Joy Division's joy distributions", subtitle = "Based on valence pulled from Spotify's Web API with spotifyr")
```

