---
title: "TidyTemplate"
date: 2022-08-02
output: html_document
editor_options: 
  chunk_output_type: inline
---

# TidyTuesday

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)

```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}
# This data loader is not working even after adding new ssh key and testing

# tt <- tt_load("2022-08-09")

tuesdata <- tidytuesdayR::tt_load(2022, week = 31)


```

```{r}
citation("tidytuesdayR")
```

```{r}
library(tidyverse)

frogs <- readr::read_csv('https://bit.ly/frog-data')

frogs |> count(Water, sort = TRUE)
```

```{r}
frogs %>% glimpse()

frogs %>% count(Subsite, sort = TRUE)
```

```{r}
frogs |> count(Type, sort = TRUE)
```

```{python}
import pandas as pd
df = pd.read_csv('https://bit.ly/frog-data')

df['Water'].value_counts()

```

```{python}
df['Type'].value_counts()

```

# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

# https://github.com/rfordatascience/tidytuesday/tree/master/data/2022/2022-08-02

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tt %>% 
  map(glimpse)

```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}



```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}


  
```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "My TidyTuesday Plot.png",
  device = "png")

```
